# Results

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(googlesheets4)
#library(blsAPI)
library(rjson) 
library(dplyr)
library(ggplot2)
library(graphics)
library("readxl")
library(tidyverse)
library("lhs")
library(splines)
library(maps)
library(mapdata)
library("viridis")
library(RColorBrewer)
library(zoo)
# library(lubridate)
library(highcharter)
```

```{r lead data, message=FALSE, warning=FALSE, echo=FALSE}
cdc = read.csv("data/cdc.csv")
cdc = cdc %>% mutate(date=as.Date(date), month=as.yearmon(format(as.Date(date), "%Y-%m")))
cdc_all = read.csv("data/cdc_all.csv", fileEncoding="UTF-8")
cdc_monthly_all_state = read.csv("data/cdc_monthly_all_state.csv", fileEncoding="UTF-8")
cpi_monthly = read.csv("data/cpi_monthly.csv", fileEncoding="UTF-8")
data_monthly = read.csv("data/data_monthly.csv", fileEncoding="UTF-8")
cdc_all = cdc_all %>% mutate(date=as.Date(date), month=as.yearmon(format(as.Date(date), "%Y-%m")))
cdc_monthly_all_state = cdc_monthly_all_state %>% mutate(month=as.yearmon(month))
cpi_monthly = cpi_monthly %>% mutate(date=as.yearmon(date))
data_monthly = data_monthly %>% mutate(month=as.yearmon(month))
```


## CPI
```{r, message=FALSE, warning=FALSE, echo=FALSE}
cpi = read.csv("data/cpidata.csv")
cpi$date <- as.yearmon(cpi$date, "%Y-%m")
```

CDC trend

```{r, fig.height = 6, message=FALSE, warning=FALSE, echo=FALSE}
cpi %>%
  ggplot( aes(x=date, y=value, group=item_name, color=item_name)) +
    geom_line() + ggtitle("CPI Time Series for All Items") + theme(axis.text.x = element_text(angle = 90),legend.text = element_text(size=5),legend.position = 'bottom', plot.title = element_text(size = 30, face = "bold"))
```


The above line chart shows the change of CPI for different items by time. Since there are so many items, we choose the ones we are most interested in to look more closely. We choose "energy","private transportation", "private transportation", "Commodities", "new vehicles", "public transportation", and "communication" for futher discuss.


Heatmap
```{r}
cor_matrix = cpi_monthly %>%  select(-date) %>% cor() %>% as.data.frame()
source('hchart _corr.R')
hchart.cor(cor_matrix)
```

According to the heat map, we can easily tell that many of the CPI items are highly and positively correlated with each other. However, `Public Transportation`, `Medical Care Commodities`, `Communication`, and `Apparel` showed white and red color in the heat map. This indicates that they are not or negatively correlated with other items. Combining with the trend graph, I know that their prices did not increase much. It is also reasonable because some of them have pricing mechanism that is not determined by the market. For instance, `Public Transportation`, `Medical Care Commodities` and `Communication` have government set prices and fixed prices. 


```{r, message=FALSE, warning=FALSE, echo=FALSE}
cpi.sub <- cpi %>% 
  filter(seriesID %in% c("SUUR0000SAT1","SUUR0000SAT", "SUUR0000SA0E","SUUR0000SAC","SUUR0000SETG","SUUR0000SETA01","SUUR0000SAE2")) 
cpi.sub %>%
  ggplot( aes(x=date, y=value, group=item_name, color=item_name)) +
    geom_line() + theme(axis.text.x = element_text(angle = 90))
```
 
We can see that the items can be divided into four groups from the above line chart. 
"energy","private transportation" and "transportation" decreases during the begining of covid-19 and then rises until now. 

"commodities" and "new vehicles" rises steadily. 

"public transportation" decreases slowly and then rise sharply during the start of 2021, then decrease sharply. 

"communication" almost no change. 
 
## Covid

### new cases and total cases change by State from 2020 Jan to 2021 Nov
We compared how new cases and total cases change from 1976 to 2020 by state

Instructions to use the graph:

Play/pause clicking the play/pause button

Navigate the motion slider by dragging the slider thumb

Navigate the motion slider by hitting the left and right arrow keys.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
cdc.month = read.csv("data/cdcmonth.csv")
cdc.month$date <- as.yearmon(cdc.month$date, "%Y-%m")
cdc.month = cdc.month  %>% group_by(statefull,date) %>% dplyr::summarise(new_case=sum(new_case),new_death=sum(new_death))

cdc_monthly_by_state = cdc %>% 
  select(month, state, new_death, new_case, conf_cases, tot_death) %>% 
  group_by(month, state) %>% 
  summarise(new_death=sum(new_death),new_case=sum(new_case), tot_cases=max(conf_cases), tot_death=max(tot_death))
cdc_monthly_by_state$statefull = state.name[match(cdc_monthly_by_state$state,state.abb)]
cdc_monthly_by_state = cdc_monthly_by_state %>% drop_na()
```


```{r}
cdcmonthcase = cdc.month %>% group_by(statefull)%>%arrange(date)%>% 
    do(item = list(
        state = first(.$statefull),
        sequence = .$new_case,
        value = first(.$new_case)
    ))%>%
    .$item
```

```{r}
highchart(type = "map") %>% 
  hc_add_series(data = cdcmonthcase,
                name = "Dem. Margin",
                mapData = usgeojson,
                joinBy = c("name","state"),
                borderWidth = 0.05,
                dataLabels = list(enabled = TRUE, format = "{point.properties.postalcode}")
                ) %>% 
  hc_colorAxis(minColor = "white", maxColor = "orange") %>%
  hc_title(text = "new cases from 2020 Apirl to 2021 Nov by state") %>%
  hc_add_theme(hc_theme_smpl()) %>%
  hc_motion(
    enabled = TRUE,
    axisLabel = "yearmonth",
    labels = as.character(sort(unique(cdc.month$date),decreasing = FALSE))
  )%>% 
  hc_legend(title = list(text = "new cases"), valueDecimals = 0, valueSuffix = "%") %>%
   hc_mapNavigation(enabled = TRUE) %>% hc_exporting(enabled = TRUE)
```

From the above motion map, we can see that the first five cases in the United States come from Illinois State, Washington State and Arizona State. At the beginning, California and New York state has the most new cases which are two of the largest States that have the most population. From Jun 2020 to Sep 2020, California, Texas and Florida States have the most new cases. Then New York comes with them to have the most new cases. Because the population for these states are really large which increases the spread of Coronavirus disease.

```{r}

cdcmonthtotal = cdc_monthly_by_state %>% group_by(statefull)%>%arrange(month)%>% 
    do(item = list(
        state = first(.$statefull),
        sequence = .$tot_cases,
        value = first(.$tot_cases)
    ))%>%
    .$item
```
```{r}
highchart(type = "map") %>% 
  hc_add_series(data = cdcmonthtotal,
                name = "Dem. Margin",
                mapData = usgeojson,
                joinBy = c("name","state"),
                borderWidth = 0.05,
                dataLabels = list(enabled = TRUE, format = "{point.properties.postalcode}")
                ) %>% 
  hc_colorAxis(minColor = "white", maxColor = "orange") %>%
  hc_title(text = "total cases from 2020 Jan to 2021 Nov by state") %>%
  hc_add_theme(hc_theme_smpl()) %>%
  hc_motion(
    enabled = TRUE,
    axisLabel = "yearmonth",
    labels = as.character(sort(unique(cdc.month$date),decreasing = FALSE))
  )%>% 
  hc_legend(title = list(text = "total cases"), valueDecimals = 0, valueSuffix = "%") %>%
   hc_mapNavigation(enabled = TRUE) %>% hc_exporting(enabled = TRUE)
```

From the motion map, we can find that at first, New York is the state with the largest total cases and then from Jul 2020 until Nov 2021, California, Texas and Flora are the three States that have the largest total cases which compatible with the trend of new cases from the previous motion map. These three States have large populations and warm weather which provides with enough condition for Coronavirus to live and spread.

### Correlation CPI V.S. Covid-19 Cases
```{r correlation}

cpi_monthly = cpi %>% select(date, value, item_name) %>% pivot_wider(id_cols = , names_from = item_name, values_from = value)
#colnames(cpi_monthly)
cdc_monthly_all_state = cdc %>% 
  select(month, new_death, new_case, tot_cases, tot_death) %>% 
  group_by(month) %>% 
  summarise(new_death=sum(new_death),new_case=sum(new_case), tot_cases=sum(tot_cases), tot_death=sum(tot_death))
#colnames(cdc_monthly_all_state)
cor_matrix = inner_join(cpi_monthly, cdc_monthly_all_state, by=c("date"="month")) %>% select(-date) %>% cor() %>% as.data.frame()

source('hchart _corr.R')
hchart.cor(cor_matrix)
```

* By looking at the pair-wise correlation between our variables(Covid data and CPI data), I am trying to use a quantitative method to find similar curves. Especially the curves that have similar trend as the total covid-19 cases. The results 


### cluster
```{r, message=FALSE, warning=FALSE, echo=FALSE}
cdc.month = read.csv("data/cdcmonth.csv")
cdc.month$date <- as.yearmon(cdc.month$date, "%Y-%m")
#cdc = read.csv("data/cdc.csv")
#cdc$date = as.Date(cdc$date)

```



```{r}
all.states = unique(cdc.month$state)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# extract top n states with the most people
top_pop_states <- function(data, n=5) {
  d = data %>% group_by(state) %>% summarize(tot_case = sum(new_case, na.rm = TRUE))
  return (top_n(d, n=n, wt=tot_case)[[1]])                          
}
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
# plot covid cases for top population states
plot_states <- function(data, states, feature, title, x=1, y=1) {
  # top_states = top_pop_states(covid, n)
  par(mfrow=c(x,y), mar=c(4,4,2,0.5)) # Setup grid, margins
  d = data %>% filter(state %in% states)
  ggplot(d, aes(x=date, y=new_case, group = state, color=state)) + geom_line() + theme(axis.text.x = element_text(angle=90)) + ggtitle(title)
    # plot(d$date, d[,feature], xlab = "Time", ylab = feature, main = paste("COVID cases in", state), pch = 20, cex = 0.5)
}
```



```{r, message=FALSE, warning=FALSE, echo=FALSE}
# returns basis of the spline to be modelled
spline.basis <- function(data, n.knots=15, degree=3) {
  X <- as.matrix(data)
  x=seq(0,1,length.out=nrow(X))
  knots = seq(0,1,length.out = n.knots-2)
  return (bs(x, knots = knots, degree = 3)[,1:(n.knots)])
}
```



```{r, message=FALSE, warning=FALSE, echo=FALSE}
# shows heatmap of the basis
B.heatmap <- function(B) {
  return (B%>%
  as.data.frame()%>%
  mutate(rw=row_number())%>%
  gather(key='key',value='value','1':'6')%>%
  ggplot(aes(x=key,y=rw,fill=value))+geom_tile()+
  labs(title='B-Spline Basis with 6 knots and 3rd-Degree Polynomials',x='Knot',y='Date'))
}

# shows multiple curves based on knots/partitions of the basis
B.basis.plot <- function(B, n.knots=10) {
  return(B%>%
  as.data.frame()%>%
  mutate(rw=row_number())%>%
  gather(key='key',value='value',`1`:paste(n.knots))%>%
  ggplot(aes(x=rw,y=value,col=key))+geom_line(aes(group=key))+
  labs(title=paste('B-Spline Basis with ', n.knots, ' knots and 3rd-Degree Polynomials', sep='')))
}
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# get spline model and return knots and spline predictions of training data
bspline.model <- function(data, dates, n.knots=6, degree=3) {
  X <- as.matrix(data)
  x=seq(0,1,length.out=nrow(X))
  knots = seq(0,1,length.out = n.knots-2)
  B = bs(x, knots = knots, degree = 3)[,1:(n.knots)]
  
  Bcoef = matrix(0,1,n.knots)
  Bcoef[1,] = solve(t(B)%*%B)%*%t(B)%*%as.matrix(X[,2])
  
  preds = (Bcoef%*%t(B))[1,]
  l = list()
  l[[1]] = Bcoef[1,]
  l[[2]] = preds
  return (l)
}
```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# get spline model and return knots and spline predictions of training data
bspline.basis <- function(data, dates, n.knots=6, degree=3) {
  X <- as.matrix(data)
  x=seq(0,1,length.out=nrow(X))
  knots = seq(0,1,length.out = n.knots-2)
  B = bs(x, knots = knots, degree = 3)[,1:(n.knots)]
  
  Bcoef = matrix(0,1,n.knots)
  Bcoef[1,] = solve(t(B)%*%B)%*%t(B)%*%as.matrix(X[,2])
  
  preds = (Bcoef%*%t(B))[1,]
  l = list()
  l[[1]] = Bcoef[1,]
  l[[2]] = B
  l[[3]] = preds
  return (l)
}
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
plot.basis.com <- function(data, states, feature, n.knots=6, degree=3, x=1, y=1) {
  # top_states = top_pop_states(covid, n)
  # par(mfrow=c(x,y), mar=c(4,4,2,0.5)) # Setup grid, margins
  data = data %>% subset(!is.na(data[, feature]))
  n = length(states)
  colors = rainbow(n.knots)
  Bcoef.mat = matrix(0,n,n.knots)
  
  for (i in 1:n) {
    plot(c(min(data$date), max(data$date)), c(min(data[,feature]), max(data[,feature])), xlab = "Time", ylab = feature, main = paste(states[i], "COVID basis demonstration"), type = "n", xaxt='n', pch = 20, cex = 0.5)
    d = data %>% filter(state == states[i]) %>% select("date", contains(feature)) %>% arrange(date)
    dates = d$date
    d$date = as.numeric(d$date)
    l = bspline.basis(d, dates, n.knots, degree)
    lines(dates, l[[3]], pch = 20, cex = 1, col="black")
    for (j in 1:n.knots) {
      l[[2]][,j] = l[[1]][j] * l[[2]][,j]
      lines(dates, l[[2]][,j], pch = 20, cex = 1, col=colors[j])
    }
    # print(paste(c("Basis", n.knots), 1:n.knots, as.character(l[[1]])))
    legend("topleft", legend=c("Estimated Spline", paste(rep("Basis", n.knots), 1:n.knots, rep("coef:", n.knots) , as.character(round(l[[1]], digits = 3)))), col=c("black", colors), lty=1, cex=0.6)
  }
  ix <- seq(1, length(dates), by=60)
  fmt <- "%b-%Y" # format for axis labels
  labs <- format(dates[ix], fmt)
  axis(side = 1, at = dates[ix], labels = labs,  cex.axis = 0.8)
}
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
# This function plots the bspline curves given the target states and parameters and returns knots for each state. 
plot.bspline <- function(data, states, feature, n.knots=6, degree=3, x=1, y=1) {
  # top_states = top_pop_states(covid, n)
  par(mfrow=c(x,y), mar=c(4,4,2,0.5)) # Setup grid, margins
  data = data %>% subset(!is.na(data[, feature]))
  n = length(states)
  colors = rainbow(n)
  Bcoef.mat = matrix(0,n,n.knots)
  
  date = as.numeric(data$date, na.rm=TRUE)
  plot(c(min(date, na.rm=TRUE), max(date, na.rm=TRUE)), c(min(data[,feature], na.rm=TRUE), max(data[,feature], na.rm=TRUE)), xlab = "Time", ylab = feature, main = "COVID cases across states", type = "n", xaxt='n', pch = 20, cex = 0.5)
  for (i in 1:n) {
    d = data %>% filter(state == states[i]) %>% select(date, contains(feature)) %>% arrange(date)
    dates = d$date
    d$date = as.numeric(d$date)
    # points(d$date, d[,feature], pch = 20, cex = 0.5)
    
    l = bspline.model(d, dates, n.knots, degree)
    lines(dates, l[[2]], pch = 20, cex = 1, col = colors[i])
    Bcoef.mat[i,] = l[[1]]
    legend("topleft", legend=states, col=colors, lty=1, cex=0.6)
    ix <- seq(1, length(dates), by=60)
    fmt <- "%b-%Y" # format for axis labels
    labs <- format(dates[ix], fmt)
    axis(side = 1, at = dates[ix], labels = labs,  cex.axis = 0.8)
  }

  return (Bcoef.mat)

}
```



```{r, message=FALSE, warning=FALSE, echo=FALSE}
state.name2abb <- function(states) {
  return (state.abb[match(states, state.name)])
}
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
# 6
Northeast = state.name2abb(c("Connecticut", "Massachusetts", "Maine", 
              "New Hampshire", "Rhode Island", "Vermont"))

# 8
Mid.Atlantic = c(state.name2abb(c("Delaware", "Maryland", 
                 "New Jersey", "New York", "Pennsylvania",
                 "Virginia", "West Virginia")),"DC")

# 9
Southeast = state.name2abb(c("Alabama", "Arkansas", "Florida",
                             "Georgia", "Louisiana", "Mississippi",
                             "North Carolina", "South Carolina", "Tennessee"))

# 13
Midwest = state.name2abb(c("Indiana", "Illinois", "Iowa", 
                           "Kansas", "Kentucky", "Michigan",
                           "Minnesota", "Missouri", "Nebraska",
                           "North Dakota", "Ohio", "South Dakota",
                           "Wisconsin"))
# 4
Southwest = state.name2abb(c("Arizona", "New Mexico", "Oklahoma", "Texas"))

# 11
West = state.name2abb(c("Alaska", "California", "Colorado", "Hawaii",
                        "Idaho", "Montana", "Nevada", "Oregon",
                        "Utah", "Washington", "Wyoming"))

region.list <- list(Northeast, Mid.Atlantic, Southeast, Midwest, Southwest, West)
names(region.list) <- c('Northeast', 'Mid.Atlantic', 'Southeast', 'Midwest', 'Southwest', 'West')
```



```{r basis plot, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
B.basis.plot(spline.basis(cdc, n.knots=10), n.knots=10)
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
mid <- function(x) {
  return ((max(x)+min(x))/2)
}
```



```{r all splines, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
Bcoef.mat = plot.bspline(cdc.month, all.states, 'new_case', n.knots=10, degree=3)
wss <- function(k) {
  kmeans(as.data.frame(Bcoef.mat), k, nstart = 10)$tot.withinss
}
# Compute and plot wss for k = 2 to k = 15
k.values <- 2:10

# extract wss for 2-15 clusters
wss_values <- map_dbl(k.values, wss)
```

```{r kmeans, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
plot(wss_values)
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
state.col = rainbow(length(all.states))
```



```{r, message=FALSE, warning=FALSE, echo=FALSE}
# plot(k.values, wss_values, type="b", pch = 19, frame = FALSE, xlab="Number of clusters K", ylab="Total within-clusters sum of squares")
map.df = map_data("state")
map.df$state = state.name2abb(str_to_title(map.df$region))
map.df$region = ifelse (map.df$state %in% Northeast, "Northeast",
                  ifelse (map.df$state %in% Mid.Atlantic, "Mid.Atlantic",
                  ifelse (map.df$state %in% Southeast, "Southeast",
                  ifelse (map.df$state %in% Midwest, "Midwest",
                  ifelse (map.df$state %in% Southwest, "Southwest",
                  ifelse (map.df$state %in% West, "West", "Other"
                          ))))))
state.centers.df = map.df %>% group_by(state) %>% summarise(long.mid = mid(long), lat.mid = mid(lat), group=mean(group))

map.state.names = geom_text(data=state.centers.df, aes(x=long.mid, y=lat.mid, group=group, label=state),size=1.5, hjust=0, vjust=0, colour = 1)
map.state.points = geom_point(data=state.centers.df, aes(x=long.mid, y=lat.mid, group=group), size = 0.5, alpha = 0.6, colour = 1)
```



```{r, message=FALSE, warning=FALSE, echo=FALSE}
sort.state.clusters <- function(clusters, sort.cluster.list) {
  clen = length(clusters)
  sort.clusters = 1:clen
  color.counter = 1
  for (c in 1:clen) {
    if (sort.cluster.list[[clusters[c]]] < 0) {
      sort.cluster.list[[clusters[c]]] = color.counter
      color.counter = color.counter + 1
    }
    sort.clusters[c] = sort.cluster.list[[clusters[c]]]
  }
  return (sort.clusters)
}
```


```{r spline map, message=FALSE, warning=FALSE, echo=FALSE}
for (i in 9) {
# Run K-means to cluster the knots of the states
  # Bcoef.mat = plot.bspline(covid.com, all.states, 'NEWCD_NORM_500', n.knots=15, degree=3)
  
  # choose the number of clusters we want and assign each state its closest cluster
  set.seed(i)

  for (k in 6) {
    cluster.num = k
    clusters = kmeans(as.data.frame(Bcoef.mat), cluster.num, nstart = 10)$cluster
    sort.cluster.list = list()
    for (j in 1:k) {
      sort.cluster.list[[j]] = -1
    }
    
    sort.clusters = sort.state.clusters(clusters, sort.cluster.list)

    states.df = data.frame(state = all.states, CLUSTER = sort.clusters)
    # write.csv(states.df, paste("./clusterings/cluster", i, ".csv", sep=""))
    
    state.cluster = list()
    for (i in 1:cluster.num) {
      state.cluster[[i]] = subset(states.df, CLUSTER==i)$state
    }
    cs = ifelse (map.df$state %in% state.cluster[[cluster.num]], cluster.num, cluster.num+1)
    for (i in (cluster.num-1):1) {
      cs = ifelse (map.df$state %in% state.cluster[[i]], i, cs)
    }
    
    map.df$cluster = cs
    #map.df$cluster = ifelse (map.df$STATE %in% state.cluster[[1]], 1,
    #                 ifelse (map.df$STATE %in% state.cluster[[2]], 2,
    #                 ifelse (map.df$STATE %in% state.cluster[[3]], 3,
    #                 ifelse (map.df$STATE %in% state.cluster[[4]], 4,
    #                 ifelse (map.df$STATE %in% state.cluster[[5]], 5,
    #                 ifelse (map.df$STATE %in% state.cluster[[6]], 6, 7
    #                        ))))))
    
    # us map on bspline knots clusters
    colors = brewer.pal(n = k, name = "Set2")
    us.bspline.map = ggplot(map.df, aes(x=long, y=lat, group = group)) + geom_polygon(colour="black", size=0.5)
    for (i in 1:cluster.num) {
      us.bspline.map = us.bspline.map + geom_polygon(fill=colors[i], data = filter(map.df, cluster == i))
    }
    print(us.bspline.map + map.state.names + map.state.points + ggtitle(paste('U.S. Map spline knots kmeans covid cases k=', k, sep='')))
  }
}
```



```{r us region map, message=FALSE, warning=FALSE, echo=FALSE}
# us map on regions
region.num = length(region.list)
region.cols = rainbow(region.num)

us.region.map = ggplot(map.df, aes(x=long, y=lat, group=group)) + 
                guides(fill=FALSE) + 
                geom_polygon(colour = "black", size = 0.5)
for (r in 1:region.num) {

  us.region.map = us.region.map + geom_polygon(fill=region.cols[r], data = filter(map.df, state %in% region.list[[r]]))
}

us.region.map + ggtitle("U.S. Map regional covid cases") + map.state.names + map.state.points
```

```{r clustered spline curves, message=FALSE, warning=FALSE, echo=FALSE}
for (i in 1:length(state.cluster)) {
  print(plot_states(cdc.month, state.cluster[[i]], "new_case", "Covid Monthly New Cases Time Series"))
}
```
To identify difference and similarities between covid case curves, we estimated a basis spline (B-spline) model for every state. Each estimated B-spline is a weighted piecewise combination of 10 polynomials, connected at “knots”. Estimated splines and weight coefficients are closely related in states with similar underlying case curves. We compared the estimated state splines by using the K-means algorithm to cluster similar sets of weight coefficients, identifying groups of states with similar covid case trajectories. In general, we can observe that the coefficients for two states in the same cluster are more similar than those in a different cluster(e.g. Florida and Texas).


