---
title: "Untitled"
author: "shuyue Xu"
date: "11/28/2021"
output: html_document
---

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
library(googlesheets4)
#library(blsAPI)
library(rjson) 
library(dplyr)
library(ggplot2)
library(graphics)
library("readxl")
library(tidyverse)
library("lhs")
library(splines)
library(maps)
library(mapdata)
library("viridis")
library(RColorBrewer)
library(zoo)
# library(lubridate)
```


```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
gs4_deauth()
data= googlesheets4::read_sheet('1fpEQ4wuuDlnfKokjKyRUrOcIwHUnIBmnBFubCUpnQ94',sheet='chained_data')
data1= googlesheets4::read_sheet('1fpEQ4wuuDlnfKokjKyRUrOcIwHUnIBmnBFubCUpnQ94',sheet='Sheet1')
chained= googlesheets4::read_sheet('1fpEQ4wuuDlnfKokjKyRUrOcIwHUnIBmnBFubCUpnQ94',sheet='chained')
```


```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
data$item_name <- chained$item_name[match(data$seriesID, chained$seriesid)]
data$item_name = as.factor(data$item_name)
data$value = as.numeric(data$value)
data$date <- paste(data$year, "-", substr(data$period,2,3), sep="")
write.csv(data,"data/cpidata.csv", row.names = FALSE)
```


```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
cpi = read.csv("data/cpidata.csv")
cpi$date <- as.yearmon(cpi$date, "%Y-%m")
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
head(cpi)
```


```{r, fig.height = 6, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
cpi %>%
  ggplot(aes(x=date, y=value, group=item_name, color=item_name)) +
    geom_line() + theme(axis.text.x = element_text(angle = 90))
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}

cpi.sub <- cpi %>% 
  filter(seriesID %in% c("SUUR0000SAT1","SUUR0000SAT", "SUUR0000SA0E","SUUR0000SAC","SUUR0000SETG","SUUR0000SETA01","SUUR0000SAE2")) 
cpi.sub %>%
  ggplot( aes(x=date, y=value, group=item_name, color=item_name)) +
    geom_line() + theme(axis.text.x = element_text(angle = 90))
'
"SUUR0000SA0E"(energy), "SUUR0000SAT1"(private transportation),"SUUR0000SAT"(transportation) decreases during the begining of covid-19 and then rises until now
 "SUUR0000SAC"(Commodities), "SUUR0000SETA01"(new vehicles) rises steadily 
 "SUUR0000SETG"(publictransportation) decreases ans then rise sharply during the start of 2021, then decrease sharply.
 "SUUR0000SAE2"(communication) no change
'
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
# plot covid cases for top population states
plot_items <- function(data, items, feature, title, x=1, y=1) {
  # top_states = top_pop_states(covid, n)
  par(mfrow=c(x,y), mar=c(4,4,2,0.5)) # Setup grid, margins
  d = data %>% filter(item_name %in% items)
  ggplot(d, aes(x=date, y=value, group = item_name, color = item_name)) + geom_line() + theme(axis.text.x = element_text(angle=90)) + ggtitle(title)
    # plot(d$date, d[,feature], xlab = "Time", ylab = feature, main = paste("COVID cases in", state), pch = 20, cex = 0.5)
}
```



```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
# get spline model and return knots and spline predictions of training data
bspline.model <- function(data, dates, n.knots=6, degree=3) {
  X <- as.matrix(data)
  x=seq(0,1,length.out=nrow(X))
  knots = seq(0,1,length.out = n.knots-2)
  B = bs(x, knots = knots, degree = 3)[,1:(n.knots)]
  
  Bcoef = matrix(0,1,n.knots)
  Bcoef[1,] = solve(t(B)%*%B)%*%t(B)%*%as.matrix(X[,2])
  
  preds = (Bcoef%*%t(B))[1,]
  l = list()
  l[[1]] = Bcoef[1,]
  l[[2]] = preds
  return (l)
}
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
# This function plots the bspline curves given the target items and parameters and returns knots for each item. 
plot.bspline <- function(data, items, feature, n.knots=6, degree=3, x=1, y=1) {
  par(mfrow=c(x,y), mar=c(4,4,2,0.5)) # Setup grid, margins
  data = data %>% subset(!is.na(data[, feature]))
  n = length(items)
  colors = rainbow(n)
  Bcoef.mat = matrix(0,n,n.knots)
  

  plot(c(min(data$date), max(data$date)), c(min(data[,feature]), max(data[,feature])), xlab = "Time", ylab = feature, main = "CPI item trends", type = "n", xaxt='n', pch = 20, cex = 0.5)
  for (i in 1:n) {
    d = data %>% filter(item_name == items[i]) %>% select(date, contains(feature)) %>% arrange(date)
    dates = d$date
    d$date = as.numeric(d$date)
    # points(d$date, d[,feature], pch = 20, cex = 0.5)
    
    l = bspline.model(d, dates, n.knots, degree)
    lines(dates, l[[2]], pch = 20, cex = 1, col = colors[i])
    Bcoef.mat[i,] = l[[1]]
    legend("topleft", legend=items, col=colors, lty=1, cex=0.2)
    ix <- seq(1, length(dates), by=60)
    fmt <- "%b-%Y" # format for axis labels
    labs <- format(dates[ix], fmt)
    axis(side = 1, at = dates[ix], labels = labs,  cex.axis = 0.8)
  }
  return (Bcoef.mat)
}
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
all.items = unique(cpi$item_name)
all.items
```


```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
Bcoef.mat = plot.bspline(cpi, all.items, 'value', n.knots=6)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
wss <- function(k) {
  kmeans(as.data.frame(Bcoef.mat), k, nstart = 5)$tot.withinss
}
# Compute and plot wss for k = 2 to k = 15
k.values <- 2:10

# extract wss for 2-15 clusters
wss_values <- map_dbl(k.values, wss)
```



```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
sort.item.clusters <- function(clusters, sort.cluster.list) {
  clen = length(clusters)
  sort.clusters = 1:clen
  color.counter = 1
  for (c in 1:clen) {
    if (sort.cluster.list[[clusters[c]]] < 0) {
      sort.cluster.list[[clusters[c]]] = color.counter
      color.counter = color.counter + 1
    }
    sort.clusters[c] = sort.cluster.list[[clusters[c]]]
  }
  return (sort.clusters)
}
```


```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
for (i in 1) {
# Run K-means to cluster the knots of the states
  
  # choose the number of clusters we want and assign each state its closest cluster
  set.seed(i)

  for (k in 6) {
    cluster.num = k
    clusters = kmeans(as.data.frame(Bcoef.mat), cluster.num, nstart = 5)$cluster
    
    sort.cluster.list = list()
    for (j in 1:k) {
      sort.cluster.list[[j]] = -1
    }
    
    sort.clusters = sort.item.clusters(clusters, sort.cluster.list)

    items.df = data.frame(item = all.items, CLUSTER = sort.clusters)
    # write.csv(states.df, paste("./clusterings/cluster", i, ".csv", sep=""))
    
    item.cluster = list()
    for (i in 1:cluster.num) {
      item.cluster[[i]] = subset(items.df, CLUSTER==i)$item
    }
  }
}
```



```{r, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
for (i in 1:length(item.cluster)) {
  print(plot_items(cpi, item.cluster[[i]], "value", "CPI Time Series"))
}
```


### bar plot
```{r}
cpi = read.csv("data/cpidata.csv")
cpi$date = as.character(cpi$date)
itemname = levels(as.factor(cpi$item_name))
```

```{r}
DF <- data.frame(item_name = character(),
                 date =character() ,
                 value = numeric(), 
                 ratebefore = numeric(), 
                 rateafter=numeric())

for (i in itemname){
  v = cpi %>% select(item_name,date,value) %>% filter(item_name == i,date == "2020-01")
  v = v$value
  b = cpi %>% select(item_name,date,value) %>% filter(item_name == i,date == "2019-01")
  b = b$value
  temp = cpi %>% select(item_name,date,value) %>% filter(item_name == i,date == "2021-10")
  t =  cpi %>% select(item_name,date,value) %>% filter(item_name == i,date == "2019-12")
  t = (t$value - b)/b
  temp$rateafter = (temp$value - v)/v
  temp$ratebefore = t
  DF = bind_rows(DF,temp)
}
```


```{r}
c = select(DF,c("item_name","ratebefore","rateafter"))
c =c %>% gather("type", "value", -item_name)
```

```{r}
ggplot(c, aes(x=item_name,y = value,fill = type))+
  geom_bar(stat="identity", position = position_dodge())+theme(axis.text.x = element_text(angle = 90))
```

From the bar plot, we can see that public transportation is the only one that decreases the CPI after the Covid-19. Because of the Covid-19, many people avoided using pubilic transportation which is crowded. The decrease of demands decreased the CPI. In the opposite, private transportation, Energy and transportation CPI increases a lot after the Covid-19 comparing with the change rates before the Covid-19. More people choose to use private transportation to reduce the chance of exposure to the crowds. We already know that energy is a necessity for private transportation. The increase of demands on these three products increases the CPI at the same time. Durable goods is the one that has negative change in CPI before the Covid-19 and then increases sharply after the Covid-19. We know that during the Covid-19, people are forced to stay at home and this may be the reason that people bought lots of durable goods to prevent emergencies.