---
title: "Untitled"
author: "shuyue Xu"
date: "11/28/2021"
output: html_document
---

```{r}
library(googlesheets4)
<<<<<<< HEAD
#library(blsAPI)
library(rjson) 
library(dplyr)
library(ggplot2)
# library(lubridate)
```


```{r}
"
gs4_deauth()
data= googlesheets4::read_sheet('1fpEQ4wuuDlnfKokjKyRUrOcIwHUnIBmnBFubCUpnQ94',sheet='chained_data')
data1= googlesheets4::read_sheet('1fpEQ4wuuDlnfKokjKyRUrOcIwHUnIBmnBFubCUpnQ94',sheet='Sheet1')
"
chained= googlesheets4::read_sheet('1fpEQ4wuuDlnfKokjKyRUrOcIwHUnIBmnBFubCUpnQ94',sheet='chained')
```


```{r}
'
data$item_name <- chained$item_name[match(data$seriesID, chained$seriesid)]
data$item_name = as.factor(data$item_name)
data$value = as.numeric(data$value)
data$date <- paste(data$year, "-", substr(data$period,2,3), sep="")
'
#write.csv(data,"data/cpidata.csv", row.names = FALSE)
```


```{r}
data = read.csv("data/cpidata.csv")
```

```{r}
data$date <- as.yearmon(data$date, "%Y-%m")
```

```{r}
=======
library(blsAPI)
library(rjson) 
library(dplyr)
library(ggplot2)
library(lubridate)
gs4_deauth()
data= googlesheets4::read_sheet('1fpEQ4wuuDlnfKokjKyRUrOcIwHUnIBmnBFubCUpnQ94',sheet='chained_data')
data1= googlesheets4::read_sheet('1fpEQ4wuuDlnfKokjKyRUrOcIwHUnIBmnBFubCUpnQ94',sheet='Sheet1')
chained= googlesheets4::read_sheet('1fpEQ4wuuDlnfKokjKyRUrOcIwHUnIBmnBFubCUpnQ94',sheet='chained')
```
```{r}
test = read.csv("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/5_OneCatSevNumOrdered.csv")
test2 = read.csv("/Users/shuyuexu/Documents/COLUMBIA/2021fall/stat5702/finalproject/cpidata.csv")
```

```{r}
data$value = as.numeric(data$value)
data$date <- paste(data$year, data$period)
head(data)
```
```{r}
data$item_name <- chained$item_name[match(data$seriesID, chained$seriesid)]
data$item_name= as.factor(data$item_name)
>>>>>>> de7cc86ba2a18db937a2254a2852bbe00231390d
head(data)
```


<<<<<<< HEAD

```{r, fig.height = 6}
=======
```{r, fig.height = 6}

>>>>>>> de7cc86ba2a18db937a2254a2852bbe00231390d
data %>%
  ggplot( aes(x=date, y=value, group=item_name, color=item_name)) +
    geom_line() + theme(axis.text.x = element_text(angle = 90))
```

```{r}
don <- data %>% 
  filter(seriesID %in% c("SUUR0000SAT1","SUUR0000SAT", "SUUR0000SA0E","SUUR0000SAC","SUUR0000SETG","SUUR0000SETA01","SUUR0000SAE2")) 
don %>%
  ggplot( aes(x=date, y=value, group=item_name, color=item_name)) +
    geom_line() + theme(axis.text.x = element_text(angle = 90))
```
 "SUUR0000SA0E"(energy), "SUUR0000SAT1"(private transportation),"SUUR0000SAT"(transportation) decreases during the begining of covid-19 and then rises until now
 "SUUR0000SAC"(Commodities), "SUUR0000SETA01"(new vehicles) rises steadily 
 "SUUR0000SETG"(publictransportation) decreases ans then rise sharply during the start of 2021, then decrease sharply.
 "SUUR0000SAE2"(communication) no change
<<<<<<< HEAD


```{r}
#test = read.csv("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/5_OneCatSevNumOrdered.csv")

```
=======
 
```{r}
don <- data %>% 
  filter(seriesID %in% c("SUUR0000SAF1","SUUR0000SAF1", "SUUR0000SAF11","SUUR0000SAF116")) 
don %>%
  ggplot( aes(x=date, y=value, group=item_name, color=item_name)) +
    geom_line() + theme(axis.text.x = element_text(angle = 90))
```
```{r}
write.csv(data,"/Users/shuyuexu/Documents/COLUMBIA/2021fall/stat5702/finalproject/cpidata.csv", row.names = FALSE)
```

>>>>>>> de7cc86ba2a18db937a2254a2852bbe00231390d
