---
title: "oil"
author: "shuyue Xu"
date: "12/14/2021"
output: html_document
---

```{r all splines, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
library(readr)
library(zoo)
library(dplyr) 
oil <- read_csv("data/oil.csv")
colnames(oil)  = c('Month','oil')
```

```{r all splines, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
oil %>%
  hchart('line', hcaes(x = 'Month', y = 'oil')) %>%
  hc_colors("grey")%>%
  hc_chart(zoomType = "x") %>%
  hc_legend(align = "right", verticalAlign = "top",layout = "horizontal",floating=TRUE,y=80,x=-100) %>%
  hc_xAxis(title = list(text = "date"),
           labels= list(style=list(fontSize='10px'),
                        rotation = -45, step = 1)) %>%
  hc_yAxis(title = list(text = "oil supply")) %>%
  hc_title(text = "<b>change of oil supply from Jan 2017 to  Sep 2021 </b>") %>%
  hc_subtitle(text = "Click on the dot to show selected month; click on a dot and drag across to zoom in") %>%
  hc_exporting(enabled = TRUE)
```


From the above line chart, we can see that before the covid-19, the supply of oil increases steadily. At the begining of Covid, from Mar 2020 to May 2020, the oil production decreases charply which shares the same trend with travel number. The decrease of needs for travel leads to the decrease of needs for energy, leads to the decrease of energy CPI and finally cause the decrease of supply. Then the production increases slowly, though there is still some decreases, finally tend to steady now. From the CPI of Energy plot and the travel number plot, we can see that both of them increases from Jan 2021 until now. The increase of needs for travel increases the needs for energy. The extra needs for energy and the shortage of energy production together makes the energy CPI increase.