# Results

```{r packages, message=FALSE, warning=FALSE, echo=FALSE}
library(googlesheets4)
library(rjson) 
library(tidyverse)
library(graphics)
library("readxl")
library("lhs")
library(splines)
library(maps)
library(mapdata)
library("viridis")
library(RColorBrewer)
library(zoo)
# library(lubridate)
library(highcharter)
```

```{r load data, message=FALSE, warning=FALSE, echo=FALSE}
cdc = read.csv("data/cdc.csv")
cdc_all = read.csv("data/cdc_all.csv", fileEncoding="UTF-8")
cdc_monthly_all_state = read.csv("data/cdc_monthly_all_state.csv", fileEncoding="UTF-8")
cpi_monthly = read.csv("data/cpi_monthly.csv", fileEncoding="UTF-8")
data_monthly = read.csv("data/data_monthly.csv", fileEncoding="UTF-8")
cdc.month = read.csv("data/cdcmonth.csv")


```

```{r simple data change, message=FALSE, warning=FALSE, echo=FALSE}
cdc = cdc %>% mutate(date=as.Date(date), month=as.yearmon(format(as.Date(date), "%Y-%m")))

cdc_all = cdc_all %>% mutate(date=as.Date(date), month=as.yearmon(format(as.Date(date), "%Y-%m")))

cdc_monthly_all_state = cdc_monthly_all_state %>% mutate(month=as.yearmon(month))

cpi_monthly = cpi_monthly %>% mutate(date=as.yearmon(date))

data_monthly = data_monthly %>% mutate(month=as.yearmon(month))

cdc.month = cdc.month %>% 
  mutate(date=as.yearmon(date, "%Y-%m")) %>% 
  group_by(statefull,date) %>% 
  dplyr::summarise(new_case=sum(new_case),new_death=sum(new_death))

cdc_monthly_by_state = cdc %>% 
  select(month, state, new_death, new_case, conf_cases, tot_death) %>% 
  group_by(month, state) %>% 
  summarise(new_death=sum(new_death),new_case=sum(new_case), tot_cases=max(conf_cases), tot_death=max(tot_death)) %>% 
  mutate(statefull=state.name[match(state,state.abb)]) %>% 
  drop_na()
```

## Covid Development in U.S.

### Cases Trend Line

#### Cumulative Total Cases

```{r}
ggplot(cdc_all,
       aes(x=date, y=conf_cases/1000)) +
  geom_area(alpha=0.8, fill="#0d2b53") + 
   scale_x_date(date_labels = "%m-%d",
               date_breaks= "4 week" 
               ) + 
  labs(x = "Date", y = "Cases Number (in thousands)", caption = "Source: CDC", 
       title = "Covid Development Trend in U.S." ,subtitle = "Cumulative Total Confirmed Cases"
       ) +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "bold", color = "grey35"),
        plot.caption = element_text(color = "grey68"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom")
```

#### Daily New Cases
```{r New Cases Trend Line, message=FALSE}
ggplot(cdc_all,
       aes(x=date, y=new_case/1000)) +
  #geom_point() + 
  geom_area(alpha=0.5, fill="#2a6998") + 
  #geom_line(stat = "smooth", method = "loess", span=0.3, size = 1.5, color="firebrick2") + 
  geom_smooth(method = "loess", span=0.3, size = 1.5, color='#95b5f0', se=F) +
  scale_x_date(date_labels = "%m-%d",
               date_breaks= "4 week" #,limits=c(as.Date('2020-03-01'), max(cdc_all$date))
               ) + 
  #scale_y_continuous(limits=c(0,15)) +
  scale_color_manual(values=c("dodgerblue3", "firebrick2")) + 
  labs(x = "Date", y = "Cases Number (in thousands)", caption = "Source: CDC", 
       title = "Covid Development Trend in U.S." ,subtitle = "Daily New Cases"
       ) +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "bold", color = "grey35"),
        plot.caption = element_text(color = "grey68"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom")

```

## Covid Development in Different States

We compared how new cases and total cases change from 1976 to 2020 by state

Instructions to use the graph:

Play/pause clicking the play/pause button

Navigate the motion slider by dragging the slider thumb

Navigate the motion slider by hitting the left and right arrow keys.

```{r}
cdcmonthcase = cdc.month %>% group_by(statefull)%>%arrange(date)%>% 
    do(item = list(
        state = first(.$statefull),
        sequence = .$new_case,
        value = first(.$new_case)
    ))%>%
    .$item

highchart(type = "map") %>% 
  hc_add_series(data = cdcmonthcase,
                name = "Dem. Margin",
                mapData = usgeojson,
                joinBy = c("name","state"),
                borderWidth = 0.05,
                dataLabels = list(enabled = TRUE, format = "{point.properties.postalcode}")
                ) %>% 
  hc_colorAxis(minColor = "white", maxColor = "orange") %>%
  hc_title(text = "new cases from 2020 Apirl to 2021 Nov by state") %>%
  hc_add_theme(hc_theme_smpl()) %>%
  hc_motion(
    enabled = TRUE,
    axisLabel = "yearmonth",
    labels = as.character(sort(unique(cdc.month$date),decreasing = FALSE))
  )%>% 
  hc_legend(title = list(text = "new cases"), valueDecimals = 0, valueSuffix = "%") %>%
   hc_mapNavigation(enabled = TRUE) %>% hc_exporting(enabled = TRUE)
```

From the above motion map, we can see that the first five cases in the United States come from Illinois State, Washington State and Arizona State. At the beginning, California and New York state has the most new cases which are two of the largest States that have the most population. From Jun 2020 to Sep 2020, California, Texas and Florida States have the most new cases. Then New York comes with them to have the most new cases. Because the population for these states are really large which increases the spread of Coronavirus disease.

```{r}
cdcmonthtotal = cdc_monthly_by_state %>% 
  group_by(statefull)%>%
  arrange(month)%>% 
    do(item = list(
        state = first(.$statefull),
        sequence = .$tot_cases,
        value = first(.$tot_cases)
    ))%>%
    .$item

highchart(type = "map") %>% 
  hc_add_series(data = cdcmonthtotal,
                name = "Dem. Margin",
                mapData = usgeojson,
                joinBy = c("name","state"),
                borderWidth = 0.05,
                dataLabels = list(enabled = TRUE, format = "{point.properties.postalcode}")
                ) %>% 
  hc_colorAxis(minColor = "white", maxColor = "orange") %>%
  hc_title(text = "total cases from 2020 Jan to 2021 Nov by state") %>%
  hc_add_theme(hc_theme_smpl()) %>%
  hc_motion(
    enabled = TRUE,
    axisLabel = "yearmonth",
    labels = as.character(sort(unique(cdc.month$date),decreasing = FALSE))
  )%>% 
  hc_legend(title = list(text = "total cases"), valueDecimals = 0, valueSuffix = "%") %>%
   hc_mapNavigation(enabled = TRUE) %>% hc_exporting(enabled = TRUE)
```

From the motion map, we can find that at first, New York is the state with the largest total cases and then from Jul 2020 until Nov 2021, California, Texas and Flora are the three States that have the largest total cases which compatible with the trend of new cases from the previous motion map. These three States have large populations and warm weather which provides with enough condition for Coronavirus to live and spread.

## Clustering States by Covid Trend
























