# Results

```{r packages, message=FALSE, warning=FALSE, echo=FALSE}
library(googlesheets4)
library(rjson) 
library(tidyverse)
library(graphics)
library("readxl")
library("lhs")
library(splines)
library(maps)
library(mapdata)
library("viridis")
library(RColorBrewer)
library(zoo)
# library(lubridate)
library(highcharter)
```

```{r load data, message=FALSE, warning=FALSE, echo=FALSE}
cdc = read.csv("data/cdc.csv")
cdc_all = read.csv("data/cdc_all.csv", fileEncoding="UTF-8")
cdc_monthly_all_state = read.csv("data/cdc_monthly_all_state.csv", fileEncoding="UTF-8")
cpi_monthly = read.csv("data/cpi_monthly.csv", fileEncoding="UTF-8")
data_monthly = read.csv("data/data_monthly.csv", fileEncoding="UTF-8")
cdc.month = read.csv("data/cdcmonth.csv")
cpi = read.csv("data/cpidata.csv")

```

```{r simple data change, message=FALSE, warning=FALSE, echo=FALSE}
cdc = cdc %>% mutate(date=as.Date(date), month=as.yearmon(format(as.Date(date), "%Y-%m")))

cdc_all = cdc_all %>% mutate(date=as.Date(date), month=as.yearmon(format(as.Date(date), "%Y-%m")))

cdc_monthly_all_state = cdc_monthly_all_state %>% mutate(month=as.yearmon(month))

cpi_monthly = cpi_monthly %>% mutate(date=as.yearmon(date))

data_monthly = data_monthly %>% mutate(month=as.yearmon(month))

cdc.month = cdc.month %>% 
  mutate(date=as.yearmon(date, "%Y-%m")) %>% 
  group_by(statefull,date) %>% 
  dplyr::summarise(new_case=sum(new_case),new_death=sum(new_death))

cdc_monthly_by_state = cdc %>% 
  select(month, state, new_death, new_case, conf_cases, tot_death) %>% 
  group_by(month, state) %>% 
  summarise(new_death=sum(new_death),new_case=sum(new_case), tot_cases=max(conf_cases), tot_death=max(tot_death)) %>% 
  mutate(statefull=state.name[match(state,state.abb)]) %>% 
  drop_na()

cpi = cpi %>% mutate(date = as.character(date))
```

## Covid Development in U.S.

### Cases Trend Line

#### Cumulative Total Cases

```{r}
ggplot(cdc_all,
       aes(x=date, y=conf_cases/1000)) +
  geom_area(alpha=0.8, fill="#0d2b53") + 
   scale_x_date(date_labels = "%m-%d",
               date_breaks= "4 week" 
               ) + 
  labs(x = "Date", y = "Cases Number (in thousands)", caption = "Source: CDC", 
       title = "Covid Development Trend in U.S." ,subtitle = "Cumulative Total Confirmed Cases"
       ) +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "bold", color = "grey35"),
        plot.caption = element_text(color = "grey68"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom")
```

#### Daily New Cases
```{r New Cases Trend Line, message=FALSE}
ggplot(cdc_all,
       aes(x=date, y=new_case/1000)) +
  #geom_point() + 
  geom_area(alpha=0.5, fill="#2a6998") + 
  #geom_line(stat = "smooth", method = "loess", span=0.3, size = 1.5, color="firebrick2") + 
  geom_smooth(method = "loess", span=0.3, size = 1.5, color='#95b5f0', se=F) +
  scale_x_date(date_labels = "%m-%d",
               date_breaks= "4 week" #,limits=c(as.Date('2020-03-01'), max(cdc_all$date))
               ) + 
  #scale_y_continuous(limits=c(0,15)) +
  scale_color_manual(values=c("dodgerblue3", "firebrick2")) + 
  labs(x = "Date", y = "Cases Number (in thousands)", caption = "Source: CDC", 
       title = "Covid Development Trend in U.S." ,subtitle = "Daily New Cases"
       ) +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(face = "bold", color = "grey35"),
        plot.caption = element_text(color = "grey68"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position="bottom")

```

## Covid Development in Different States

We compared how new cases and total cases change from 1976 to 2020 by state

Instructions to use the graph:

Play/pause clicking the play/pause button

Navigate the motion slider by dragging the slider thumb

Navigate the motion slider by hitting the left and right arrow keys.

```{r}
cdcmonthcase = cdc.month %>% group_by(statefull)%>%arrange(date)%>% 
    do(item = list(
        state = first(.$statefull),
        sequence = .$new_case,
        value = first(.$new_case)
    ))%>%
    .$item

highchart(type = "map") %>% 
  hc_add_series(data = cdcmonthcase,
                name = "Dem. Margin",
                mapData = usgeojson,
                joinBy = c("name","state"),
                borderWidth = 0.05,
                dataLabels = list(enabled = TRUE, format = "{point.properties.postalcode}")
                ) %>% 
  hc_colorAxis(minColor = "white", maxColor = "orange") %>%
  hc_title(text = "new cases from 2020 Apirl to 2021 Nov by state") %>%
  hc_add_theme(hc_theme_smpl()) %>%
  hc_motion(
    enabled = TRUE,
    axisLabel = "yearmonth",
    labels = as.character(sort(unique(cdc.month$date),decreasing = FALSE))
  )%>% 
  hc_legend(title = list(text = "new cases"), valueDecimals = 0, valueSuffix = "%") %>%
   hc_mapNavigation(enabled = TRUE) %>% hc_exporting(enabled = TRUE)
```

From the above motion map, we can see that the first five cases in the United States come from Illinois State, Washington State and Arizona State. At the beginning, California and New York state has the most new cases which are two of the largest States that have the most population. From Jun 2020 to Sep 2020, California, Texas and Florida States have the most new cases. Then New York comes with them to have the most new cases. Because the population for these states are really large which increases the spread of Coronavirus disease.

```{r}
cdcmonthtotal = cdc_monthly_by_state %>% 
  group_by(statefull)%>%
  arrange(month)%>% 
    do(item = list(
        state = first(.$statefull),
        sequence = .$tot_cases,
        value = first(.$tot_cases)
    ))%>%
    .$item

highchart(type = "map") %>% 
  hc_add_series(data = cdcmonthtotal,
                name = "Dem. Margin",
                mapData = usgeojson,
                joinBy = c("name","state"),
                borderWidth = 0.05,
                dataLabels = list(enabled = TRUE, format = "{point.properties.postalcode}")
                ) %>% 
  hc_colorAxis(minColor = "white", maxColor = "orange") %>%
  hc_title(text = "total cases from 2020 Jan to 2021 Nov by state") %>%
  hc_add_theme(hc_theme_smpl()) %>%
  hc_motion(
    enabled = TRUE,
    axisLabel = "yearmonth",
    labels = as.character(sort(unique(cdc.month$date),decreasing = FALSE))
  )%>% 
  hc_legend(title = list(text = "total cases"), valueDecimals = 0, valueSuffix = "%") %>%
   hc_mapNavigation(enabled = TRUE) %>% hc_exporting(enabled = TRUE)
```

From the motion map, we can find that at first, New York is the state with the largest total cases and then from Jul 2020 until Nov 2021, California, Texas and Flora are the three States that have the largest total cases which compatible with the trend of new cases from the previous motion map. These three States have large populations and warm weather which provides with enough condition for Coronavirus to live and spread.

## Clustering States by Covid Trend

```{r basis plot, message=FALSE, warning=FALSE, echo=FALSE, results='hide', fig.show='hide'}
Bcoef.mat = plot.bspline(cdc.month, all.states, 'new_case', n.knots=10, degree=3)
source('alvin-clustering.r')
B.basis.plot(spline.basis(cdc, n.knots=10), n.knots=10)
```


```{r spline map, message=FALSE, warning=FALSE, echo=FALSE}
for (i in 9) {
# Run K-means to cluster the knots of the states
  # Bcoef.mat = plot.bspline(covid.com, all.states, 'NEWCD_NORM_500', n.knots=15, degree=3)
  
  # choose the number of clusters we want and assign each state its closest cluster
  set.seed(i)

  for (k in 6) {
    cluster.num = k
    clusters = kmeans(as.data.frame(Bcoef.mat), cluster.num, nstart = 10)$cluster
    sort.cluster.list = list()
    for (j in 1:k) {
      sort.cluster.list[[j]] = -1
    }
    
    sort.clusters = sort.state.clusters(clusters, sort.cluster.list)

    states.df = data.frame(state = all.states, CLUSTER = sort.clusters)
    # write.csv(states.df, paste("./clusterings/cluster", i, ".csv", sep=""))
    
    state.cluster = list()
    for (i in 1:cluster.num) {
      state.cluster[[i]] = subset(states.df, CLUSTER==i)$state
    }
    cs = ifelse (map.df$state %in% state.cluster[[cluster.num]], cluster.num, cluster.num+1)
    for (i in (cluster.num-1):1) {
      cs = ifelse (map.df$state %in% state.cluster[[i]], i, cs)
    }
    
    map.df$cluster = cs
    #map.df$cluster = ifelse (map.df$STATE %in% state.cluster[[1]], 1,
    #                 ifelse (map.df$STATE %in% state.cluster[[2]], 2,
    #                 ifelse (map.df$STATE %in% state.cluster[[3]], 3,
    #                 ifelse (map.df$STATE %in% state.cluster[[4]], 4,
    #                 ifelse (map.df$STATE %in% state.cluster[[5]], 5,
    #                 ifelse (map.df$STATE %in% state.cluster[[6]], 6, 7
    #                        ))))))
    
    # us map on bspline knots clusters
    colors = brewer.pal(n = k, name = "Set2")
    us.bspline.map = ggplot(map.df, aes(x=long, y=lat, group = group)) + geom_polygon(colour="black", size=0.5)
    for (i in 1:cluster.num) {
      us.bspline.map = us.bspline.map + geom_polygon(fill=colors[i], data = filter(map.df, cluster == i))
    }
    print(us.bspline.map + map.state.names + map.state.points + ggtitle(paste('U.S. Map spline knots kmeans covid cases k=', k, sep='')))
  }
}
```
```{r clustered spline curves, message=FALSE, warning=FALSE, echo=FALSE}
for (i in 1:length(state.cluster)) {
  print(plot_states(cdc.month, state.cluster[[i]], "new_case", "Covid Monthly New Cases Time Series"))
}
```

## Consumer Price Index Development

### Chang of Living Expenses by Items

```{r}
itemname = levels(as.factor(cpi$item_name))
DF <- data.frame(item_name = character(),
                 date =character() ,
                 value = numeric(), 
                 ratebefore = numeric(), 
                 rateafter=numeric())

for (i in itemname){
  v = cpi %>% select(item_name,date,value) %>% filter(item_name == i,date == "2020-01")
  v = v$value
  b = cpi %>% select(item_name,date,value) %>% filter(item_name == i,date == "2019-01")
  b = b$value
  temp = cpi %>% select(item_name,date,value) %>% filter(item_name == i,date == "2021-10")
  t =  cpi %>% select(item_name,date,value) %>% filter(item_name == i,date == "2019-12")
  t = (t$value - b)/b
  temp$rateafter = (temp$value - v)/v
  temp$ratebefore = t
  DF = bind_rows(DF,temp)
}
c = select(DF,c("item_name","ratebefore","rateafter"))
c =c %>% gather("type", "value", -item_name)
```

```{r}
ggplot(c, aes(x=item_name,y = value,fill = type))+
  geom_bar(stat="identity", position = position_dodge())+theme(axis.text.x = element_text(angle = 90))
```

From the bar plot, we can see that public transportation is the only one that decreases the CPI after the Covid-19. Because of the Covid-19, many people avoided using pubilic transportation which is crowded. The decrease of demands decreased the CPI. In the opposite, private transportation, Energy and transportation CPI increases a lot after the Covid-19 comparing with the change rates before the Covid-19. More people choose to use private transportation to reduce the chance of exposure to the crowds. We already know that energy is a necessity for private transportation. The increase of demands on these three products increases the CPI at the same time. Durable goods is the one that has negative change in CPI before the Covid-19 and then increases sharply after the Covid-19. We know that during the Covid-19, people are forced to stay at home and this may be the reason that people bought lots of durable goods to prevent emergencies.

### Correlation Between Consumer Items

```{r}
cor_matrix = cpi_monthly %>%  select(-date) %>% cor() %>% as.data.frame()
source('hchart _corr.R')
hchart.cor(cor_matrix)
```

According to the heat map, we can easily tell that many of the CPI items are highly and positively correlated with each other. However, `Public Transportation`, `Medical Care Commodities`, `Communication`, and `Apparel` showed white and red color in the heat map. This indicates that they are not or negatively correlated with other items. Combining with the trend graph, I know that their prices did not increase much. It is also reasonable because some of them have pricing mechanism that is not determined by the market. For instance, `Public Transportation`, `Medical Care Commodities` and `Communication` have government set prices and fixed prices.

Industrial resources are highly correlated with each other. `Energy`,  `Commodities`, `Transportation

Therefore, we grouped similar CPI items together and will show their similarities in the next part. 

### Group CPI Items

The groups we created for CPI items are as follows:
1. Price not sensitive to other items: `Public Transportation`, `Medical Care Commodities`, `Communication`, and `Apparel`
2. Energy related items: 
3. 












